---
title: Regression Project
author: Alim Ray
date: 10/25/2015
---


#Executive Summary
Using the mtcars data included in RStudio, I have been asked to answer the folowing questions.
Is an automatic or manual transmission better for MPG?
Quantify the MPG difference between automatic and manual transmissions?

#Setting up data
Reading mtcars data and adding cylinder and transmission factors.  Making transmission a factor variable will allow me to determine the marginal differences between manual and automatic transmissions when creating a linear regression.
```{r setup, echo = F}
library(dplyr)
library(ggplot2)
library(GGally)
data(mtcars)

setupData <- function() {
	mtcars$carnames <- rownames(mtcars)
	mycars <- tbl_df(mtcars)
	mycars$cylType <- mycars$cyl
	mycars$cylType <- ifelse( mycars$cyl == 4, 0, 3)
	mycars$cylType <- ifelse( mycars$cyl == 6, 1, mycars$cylType)
	mycars$cylType <- ifelse( mycars$cyl == 8, 2, mycars$cylType)
	mycars$cylType <- factor(mycars$cylType, levels = c('0', '1','2'), labels = c('4cyl', '6cyl', '8cyl'))
	mycars$am <- factor(mycars$am, levels = c('0', '1'), labels = c('Automatic', 'Manual') )
	mycars
}

cars <- setupData()

```


#Expolaratory Analysis
```{r quickmodels, echo = F}
simpleMod <- lm( mpg ~ am, mtcars)
```
Looking at the linear regression between mpg and transmissions only yields the following model.

Intercept = `r simpleMod$coefficients[1]`

Slope = `r simpleMod$coefficients[2]`

This shows a positive correlation before adjustments.  I then looked at number of cylinders and weight because I believed they would effect the model.  Below is a pairs plot of mpg, transmission, number of cylinders, and weight.

```{r pairsMPG_AM_WT, echo = F, fig.width=10, fig.height=10, fig.align='center'}

pairsGraph <- function( mycars ) {
	ggpairs( mycars, columns = c(1,6,13,9), lower = list( continuous = 'smooth'), diag = list( discrete = 'bar'), title =  "mpg, transmisison, num of cylinders, weight" )
}

pairsGraph(cars)
```

##Cylinders and Weight
Cell 1,3 of this graph is a box and whisker plot which shows that as number of cylinders increase, mpg decreases.  Cell 2,1 also shows that as weight increases, mpg decreases.  I checked the correlation between weight and cylinders and got `r cor(cars$wt, cars$cyl)`.  Since these data are highly correlated, I decided to remove weight from the model.

##Displacement and Horsepower
Engine displacement is the volume	of an engine's cylinders.  This is directly correlated to the number of cylinders (correlation = `r cor(cars$disp, cars$cyl)`) and was removed from the model.  Horsepower is also a function of cylinders (correlation = `r cor(cars$hp, cars$cyl)`) and was excluded from the model.


#Elimination of variables via ANOVA
Using the rest of the variables (and transmission + number of cylinders), I created five linear models that I tested using ANOVA.

```{r anova}
everything <- lm( mpg ~ am + drat + cylType, cars)
everything2 <- lm( mpg ~ am + drat + cylType + qsec, cars)
everything3 <- lm( mpg ~ am + drat + cylType + qsec + vs, cars)
everything4 <- lm( mpg ~ am + drat + cylType + qsec + vs + gear, cars)
everything5 <- lm( mpg ~ am + drat + cylType + qsec + vs + gear + carb, cars)
anova(everything, everything2, everything3, everything4, everything5)
```

The qsec, v/s, and number of gears are shown to have high P-values (all > 0.4).  Removing these fields, the final model is transmission, rear axle ratio, number of cylinders, and number of carburetors.

#Conclusion
```{r conclusion}
finalM <- lm( mpg ~ am + drat + cylType + carb, cars)
confInt <- summary(finalM)$coefficients[2,1] + c(-1,1) * summary(finalM)$coefficients[2,2] * qt(0.975, df = finalM$df)
```
Using the final model the coefficient for manual transmission is `r finalM$coefficients[2]`.  Since this is positive, the effect of manual transmission on miles per gallon is positive.  `r finalM$coefficents[2]` with a confidence interval of `r confInt[1]` to `r confInt[2]` is also the average mpg increase when using a manual transmission.
